Jack Antonoff-ification of pop music

### 2023-04-04 outputs/paper/paper.qmd
more qmd edits

---
title: "Jack Antonoff-ification of pop music"
subtitle: "My subtitle if needed"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: LINK."
date: today
abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API. Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an "Antonoff-y" song, and assess whether or not the audio features defined by Spotify can capture this. I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me. Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
format: html
number-sections: true
bibliography: references.bib
---

#quarto references itself. wrapping in here:here goes in project 
knitr::include_graphics(here::here("inputs/data/apod_20230227.jpg"))

```{r}
#| label: fig-energy
#| fig-cap: Energy Variable
#| echo: false

library(lubridate)
df <- readRDS(here::here("inputs/data/df.rds"))
df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

@@

To find an answer to this question, I turned to statistics, people who know more than me about describing music, and the overlord of engineered virality itself -- the Spotify API.

@@

@@

\[\[He ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]\]

@@

Taste is subjective. People who can play instruments tend to be better at identifying and describing sounds. Cultural criticism is a hall of smoke and mirrors. Antonoff is everywhere.

@@

#| echo: false
#| warning: false

@@

```{r}
library(kableExtra)
df |>
  knitr::kable()

@@

>>>>

---
title: "Jack Antonoff-ification of pop music"
subtitle: "My subtitle if needed"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: LINK."
date: today
abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API. Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an Antonoff-y song, and assess whether or not the audio features defined by Spotify can capture this. I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me. There will be a Bayesian inference model. It will be a logistic regression. We do not know whether or not it will be good. Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

To find an answer to this question, I turned to statistics, people who know more than me about describing music, and the overlord of engineered virality itself -- the Spotify API. There will be a Bayesian inference model. It will be a logistic regression. We do not know whether or not it will be good.

@@

![](images/Screen%20Shot%202023-04-03%20at%209.31.05%20PM.png)

@@

\[\[He told me that he didn't actually refute the premise of my interview, and ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]\]

@@

TLDR of the Gamman interview is: Taste is subjective. People who can play instruments tend to be better at identifying and describing sounds. Cultural criticism is a hall of smoke and mirrors. Antonoff is everywhere.

@@

#| echo: false
#| warning: false
library(dplyr)

@@

```{r}
#| echo: false
#| warning: false

#library(kableExtra)
#df |>
 # knitr::kable()

@@

# "EDA"

```{r}
#| label: fig-energy
#| fig-cap: Energy Variable
#| echo: false
# 
# library(lubridate)
# df <- read.csv(here::here("inputs/data/df.csv"))
# df %>%
#   filter(album_release_date_precision == "day") %>%
#   mutate(album_release_date = ymd(album_release_date)) %>%
#   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
#   geom_point() +
#   theme_minimal() +
#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
#   labs(color = "Has 'y' in jack?") +
#   facet_wrap(~ artist_name, ncol = 3)

# df %>%
#   filter(album_release_date_precision == "day") %>%
#   mutate(album_release_date = ymd(album_release_date)) %>%
#   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
#   geom_point() +
#   theme_minimal() +
#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
#   labs(color = "Has 'y' in jack?") +
#   facet_wrap(~ artist_name, ncol = 3)

# df %>%
#   filter(album_release_date_precision == "day") %>%
#   mutate(album_release_date = ymd(album_release_date)) %>%
#   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
#   geom_point() +
#   theme_minimal() +
#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
#   labs(color = "Has 'y' in jack?") +
#   facet_wrap(~ artist_name, ncol = 3)

# df %>%
#   filter(album_release_date_precision == "day") %>%
#   mutate(album_release_date = ymd(album_release_date)) %>%
#   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
#   geom_point() +
#   theme_minimal() +
#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
#   labs(color = "Has 'y' in jack?") +
#   facet_wrap(~ artist_name, ncol = 3)

# 
# df %>%
#   filter(album_release_date_precision == "day") %>%
#   mutate(album_release_date = ymd(album_release_date)) %>%
#   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
#   geom_point() +
#   theme_minimal() +
#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
#   labs(color = "Has 'y' in jack?") +
#   facet_wrap(~ artist_name, ncol = 3)

# 
# df %>%
#   filter(album_release_date_precision == "day") %>%
#   mutate(album_release_date = ymd(album_release_date)) %>%
#   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
#   geom_point() +
#   theme_minimal() +
#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
#   labs(color = "Has 'y' in jack?") +
#   facet_wrap(~ artist_name, ncol = 3)
